\LARGE\singlespacing{\textbf{S1 Appendix} \\ \medskip Estimating \emph{Mycobacterium} tuberculosis transmission in a South African clinic: Spatiotemporal model based on person movements}

### 2024-11-19 doc/paper/supplements.tex
Updates from Overleaf

\renewcommand{\thesection}{Text \arabic{section}}
\usepackage{titlesec}
%\titleformat{\section}{\normalfont\Large\bfseries}{Text \thesection.~#1}{1em}{}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}

@@

\title{\LARGE\singlespacing{\textbf{Supplementary Information} \\ \medskip
Estimating \emph{Mycobacterium} tuberculosis transmission in a South African clinic: Spatiotemporal model based on person movements}}

@@

Person-tracking data required processing because many tracks were interrupted, for example, if the person i)~moved outside the range of the video sensors and then re-entered, ii)~bowed heavily downwards, or iii)~was temporarily hidden behind another person or obstacle. In addition, moving objects such as opening doors may have been recognized as a person, introducing many extremely short tracks (total duration of less than 10\,s and distance less than 2\,m), which were removed.

@@

The $AER$ is not directly observed but can be estimated from indoor CO$_2$ levels, room occupancy, and volume. Under steady-state conditions (\ie CO$_2$ levels reaching a steady-state mixing ratio), the $AER$ can be computed as
\begin{align}
    AER = \frac{6\cdot10^4 \cdot n \cdot G}{V\cdot(CO^{\text{I}}-CO^{\text{O}})},
\end{align}
where $n$ is the number of people in the airspace, $G$ is the average CO$_2$ generation rate per person, $CO^{\text{I}}$ is the CO$_2$ mixing ratio in indoor and $CO^{\text{O}}$ is the mixing ratio in outdoor air\cite{Batterman2017IJERPH}. However, in a primary care clinic, room occupancy varies continuously and thus it may be difficult to determine the steady-state CO$_2$ mixing ratio. Therefore, we determine $AER$ using a transient mass balance model of the form\cite{Batterman2017IJERPH}
\begin{align}
    CO_{t+1}^{\text{I}} = \frac{6\cdot10^4 \cdot n_t + G}{Q} \cdot \left(1 - \exp(-Q/V \Delta t)\right) + (CO_t^{\text{I}}-CO^{\text{O}}) \cdot \exp(-Q/V \Delta t) + CO^{\text{O}},
\end{align}
where $Q$ is the outdoor air supply rate (in m$^3$\,s$^{-1}$ $\cdot \frac{1}{3600}$ ) and $AER = Q/V$. We numerically solve the transient mass balance model with the limited memory Broyden–Fletcher–Goldfarb–Shanno algorithm (L-BFGS), a quasi-Newton method, as implemented in the statistical software R\cite{Byrd1995SIAM}. The optimal solution is the outdoor air supply rate $Q$ that minimizes the root mean square error. We also follow recommendations to fit the outdoor CO$_2$ mixing ratio with reasonable constraints on $CO^{\text{O}} = [300,600]$\,ppm\cite{Batterman2017IJERPH}. We can compute $AER$ over different time windows, hence the index in $AER_t$.

@@

We illustrate our modelling approach with an example. Consider a room with a volume $V$=150\,m$^3$ (length = 10\,m, breadth = 5\,m, height = 3\,m), which we rasterize into a grid of $C = 40 \cdot 20 = 800$\,cells with an area of (0.25\,m)$^2$. We assume that quanta generation is confined to the cell where the infectious individual is located as well as the first neighbouring cells. We monitor the room from 8\,am to 12\,am and consider one infectious individual generating quanta continuously for one hour between 8\,am and 9\,am in the middle of the room (cell (20,10)), which means that quanta is generated in cells $(x_c = 20\pm1, y_c = 10\pm1)$. We assume a quanta generation rate of $q = 2$\,quanta h$^{-1}$ and a bacterial inactivation rate of $\lambda = 0$\,quanta h$^{-1}$. We further assume a constant outdoor air exchange rate of $AER = 1$\,air changes h$^{-1}$, translating into a quanta removal rate of 1\,quanta h$^{-1}$ and corresponding to a diffusion constant of 0.015\,m$^2$ s$^{-1}$.

@@

We divide the clinic into three spaces: the waiting room (length = 10.55\,m $\times$ breadth = 5.50\,m $\times$ height = 3.00\,m), the corridor (length = 12.45\,m $\times$ breadth = 2.20\,m $\times$ height = 2.50\,m), and the TB room (length = 4.75\,m $\times$ breadth = 3.50\,m $\times$ height = 3.00\,m). We rasterize each room into a grid of cubic cells with a squared area of (0.25\,m)$^2$, corresponding to a diagonal length of $\sqrt{0.25^2 + 0.25^2} = 0.35$\,m.

Although the door between the waiting room and corridor was usually not closed, we consider these rooms as separate spaces to facilitate modelling and because we observed differences in room occupancy and time-varying CO$_2$ levels. Nevertheless, we acknowledge the possibility for aerosolized \emph{Mtb} (quanta) to spread from the waiting room to the corridor, and vice versa, which is not considered in our simulations.

@@

The bacterial inactivation rate for \emph{Mtb} has not been precisely estimated. Loudon (1969)\cite{Loudon1969AMRRD} report a half-life for aerosolized TB bacilli of 6\,h ($\lambda$ = 0.12). By contrast, Lever (2000)\cite{Lever2000LettersAppliedMicrobio} reported a half-life of just 5\,min ($\lambda$ = 8.3). The difference could be explained by differences in humidity levels during the study\cite{Lever2000LettersAppliedMicrobio}: Loudon (1969) and Lever (2000) studied survival in an environment with relative humidity of about 50\% and 70\%, respectively. Relative humidity in our study is closer to 50\%, thus the estimate by Loudon (1969) may be more suitable for our study. Furthermore, Gannon (2007)\cite{Gannon2007ResVetSci} report a half-life of 1.5\,h ($\lambda$ = 0.46) at relative humidity of over 75\%, yet the estimate is for \emph{Mtb bovis}, and Klein (2014)\cite{Klein2014IJMyco} observe a 80\% reduction within 30\,min ($\lambda$ = 3.2), which is closer to the estimate by Lever (2000). Considering the variation in these estimates, we model $\lambda$ with a $\mathrm{Lognormal}(\mu = \log 1, \sigma = 1)$ distribution, with a median rate of 1\,quanta h$^{-1}$ (95\%-CrI 0.1\,quanta h$^{-1}$\,$-$\,7.1\,quanta h$^{-1}$). The distribution is shown in \Cref{fig:lambda-distribution} along with the estimates from the literature.

@@

\begin{figure}[!htpb]
    \centering
    \includegraphics{results/data/co2-levels-over-time.png}
    \caption{CO$_2$ levels by clinic area over time for each study day.}
    \label{fig:co2-levels}
\end{figure}

@@

The generation, dispersion, and survival of infectious particles in the air is subject to considerable uncertainty. Therefore, Wells\cite{Wells1955} introduced the concept of a ``quantum'' (infectious dose) to capture the stochastic nature of airborne transmission. A higher concentration of infectious quanta corresponds to a higher probability for an individual to get infected. The probability of infection is modeled with a Poisson relation and the Wells-Riley model considers multiple determinants for the generation and survival of infectious quanta. Nevertheless, transmission events are difficult to predict as the occurrence depends on multiple other factors that can broadly be categorized into (1)~environmental factors, (2)~physicochemical properties of the pathogen, and the (3)~infectiousness of the source and susceptibility of the host.

>>>>

\renewcommand{\thesection}{Text \arabic{section}}
\usepackage{titlesec}
%\titleformat{\section}{\normalfont\Large\bfseries}{Text \thesection.~#1}{1em}{}
\renewcommand{\thefigure}{\Alph{figure}}
\renewcommand{\thetable}{\Alph{table}}

@@

\title{\LARGE\singlespacing{\textbf{S1 Appendix} \\ \medskip
Estimating \emph{Mycobacterium} tuberculosis transmission in a South African clinic: Spatiotemporal model based on person movements}}

@@

Person-tracking data required processing because many tracks were interrupted, for example, if the person i)~moved outside the range of the video sensors and then re-entered, ii)~bowed heavily downwards, or iii)~was temporarily hidden behind another person or obstacle. In addition, moving objects such as opening doors may have been recognised as a person, introducing many extremely short tracks (total duration of less than 10\,s and distance less than 2\,m), which were removed.

@@

The $AER$ is not directly observed but can be estimated from indoor CO$_2$ levels, room occupancy, and volume. Under steady-state conditions (\ie CO$_2$ levels reaching a steady-state mixing ratio), the $AER$ can be computed as
\begin{align}
    AER = \frac{6\cdot10^4 \cdot n \cdot G}{V\cdot(CO^{\text{I}}-CO^{\text{O}})},
\end{align}
where $n$ is the number of people in the airspace, $G$ is the average CO$_2$ generation rate per person, $CO^{\text{I}}$ is the CO$_2$ mixing ratio in indoor and $CO^{\text{O}}$ is the mixing ratio in outdoor air\cite{Batterman2017IJERPH}. However, in a primary care clinic, room occupancy varies continuously and thus it may be difficult to determine the steady-state CO$_2$ mixing ratio. Therefore, we determine $AER$ using a transient mass balance model of the form\cite{Batterman2017IJERPH}
\begin{align}
    CO_{t+1}^{\text{I}} = \frac{6\cdot10^4 \cdot n_t + G}{Q} \cdot \left(1 - \exp(-Q/V \Delta t)\right) + (CO_t^{\text{I}}-CO^{\text{O}}) \cdot \exp(-Q/V \Delta t) + CO^{\text{O}},
\end{align}
where $Q$ is the outdoor air supply rate (in m$^3$\,s$^{-1}$ $\cdot \frac{1}{3600}$ ) and $AER = Q/V$. We numerically solve the transient mass balance model with the limited memory Broyden–Fletcher–Goldfarb–Shanno algorithm (L-BFGS), a quasi-Newton method, as implemented in the statistical software R\cite{Byrd1995SIAM}. The optimal solution is the outdoor air supply rate $Q$ that minimises the root mean square error. We also follow recommendations to fit the outdoor CO$_2$ mixing ratio with reasonable constraints on $CO^{\text{O}} = [300,600]$\,ppm\cite{Batterman2017IJERPH}. We can compute $AER$ over different time windows, hence the index in $AER_t$.

@@

We illustrate our modelling approach with an example. Consider a room with a volume $V$=150\,m$^3$ (length = 10\,m, breadth = 5\,m, height = 3\,m), which we rasterise into a grid of $C = 40 \cdot 20 = 800$\,cells with an area of (0.25\,m)$^2$. We assume that quanta generation is confined to the cell where the infectious individual is located as well as the first neighbouring cells. We monitor the room from 8\,am to 12\,am and consider one infectious individual generating quanta continuously for one hour between 8\,am and 9\,am in the middle of the room (cell (20,10)), which means that quanta is generated in cells $(x_c = 20\pm1, y_c = 10\pm1)$. We assume a quanta generation rate of $q = 2$\,quanta h$^{-1}$ and a bacterial inactivation rate of $\lambda = 0$\,quanta h$^{-1}$. We further assume a constant outdoor air exchange rate of $AER = 1$\,air changes h$^{-1}$, translating into a quanta removal rate of 1\,quanta h$^{-1}$ and corresponding to a diffusion constant of 0.015\,m$^2$ s$^{-1}$.

@@

We divide the clinic into three spaces: the waiting room (length = 10.55\,m $\times$ breadth = 5.50\,m $\times$ height = 3.00\,m), the corridor (length = 12.45\,m $\times$ breadth = 2.20\,m $\times$ height = 2.50\,m), and the TB room (length = 4.75\,m $\times$ breadth = 3.50\,m $\times$ height = 3.00\,m). We rasterise each room into a grid of cubic cells with a squared area of (0.25\,m)$^2$, corresponding to a diagonal length of $\sqrt{0.25^2 + 0.25^2} = 0.35$\,m.

Although the door between the waiting room and corridor was usually not closed, we consider these rooms as separate spaces to facilitate modelling and because we observed differences in room occupancy and time-varying CO$_2$ levels. Nevertheless, we acknowledge the possibility for aerosolised \emph{Mtb} (quanta) to spread from the waiting room to the corridor, and vice versa, which is not considered in our simulations.

@@

The bacterial inactivation rate for \emph{Mtb} has not been precisely estimated. Loudon (1969)\cite{Loudon1969AMRRD} report a half-life for aerosolised TB bacilli of 6\,h ($\lambda$ = 0.12). By contrast, Lever (2000)\cite{Lever2000LettersAppliedMicrobio} reported a half-life of just 5\,min ($\lambda$ = 8.3). The difference could be explained by differences in humidity levels during the study\cite{Lever2000LettersAppliedMicrobio}: Loudon (1969) and Lever (2000) studied survival in an environment with relative humidity of about 50\% and 70\%, respectively. Relative humidity in our study is closer to 50\%, thus the estimate by Loudon (1969) may be more suitable for our study. Furthermore, Gannon (2007)\cite{Gannon2007ResVetSci} report a half-life of 1.5\,h ($\lambda$ = 0.46) at relative humidity of over 75\%, yet the estimate is for \emph{Mtb bovis}, and Klein (2014)\cite{Klein2014IJMyco} observe a 80\% reduction within 30\,min ($\lambda$ = 3.2), which is closer to the estimate by Lever (2000). Considering the variation in these estimates, we model $\lambda$ with a $\mathrm{Lognormal}(\mu = \log 1, \sigma = 1)$ distribution, with a median rate of 1\,quanta h$^{-1}$ (95\%-CrI 0.1\,quanta h$^{-1}$\,$-$\,7.1\,quanta h$^{-1}$). The distribution is shown in \Cref{fig:lambda-distribution} along with the estimates from the literature.

@@

\begin{figure}[!htpb]
    \centering
    \includegraphics{results/data/co2-levels-over-time.png}
    \caption{CO$_2$ levels by clinic area over time for each study day. Raw data at 1~min intervals averaged over a ten minute time window to reduce noise.}
    \label{fig:co2-levels}
\end{figure}

\begin{figure}[!htpb]
    \centering
    \includegraphics{results/data/no-people-over-time-by-day.png}
    \caption{Number of people by clinic area over time for each study day. Raw data at 1~second intervals  averaged over a ten minute time window to reduce noise.}
    \label{fig:people-over-time-by-day}
\end{figure}

@@

The generation, dispersion, and survival of infectious particles in the air is subject to considerable uncertainty. Therefore, Wells\cite{Wells1955} introduced the concept of a ``quantum'' (infectious dose) to capture the stochastic nature of airborne transmission. A higher concentration of infectious quanta corresponds to a higher probability for an individual to get infected. The probability of infection is modeled with a Poisson relation and the Wells-Riley model considers multiple determinants for the generation and survival of infectious quanta. Nevertheless, transmission events are difficult to predict as the occurrence depends on multiple other factors that can broadly be categorised into (1)~environmental factors, (2)~physicochemical properties of the pathogen, and the (3)~infectiousness of the source and susceptibility of the host.
